---
title: "Bujarski Final Project"
author: "Kevin Bujarski"
date: "March 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Point Pattern and Area Analysis of Asheville, NC Airbnb listings  

This project will look at Airbnb listings for Asheville, NC.  We will apply spatial interpolation methods and multiple linear regression models to deterine an estimated price of a room type.

### Data Prep

```{r import}
library(sf)
library(sp)
library(maptools)
library(tidyverse)
library(ggmap)
library(jsonlite)
library(rgdal)
library(GISTools)
library(tmap)

asheville <- read.csv(file = "http://data.insideairbnb.com/united-states/nc/asheville/2020-02-29/visualisations/listings.csv")

asheville_polygons <- st_read("http://data.insideairbnb.com/united-states/nc/asheville/2020-02-29/visualisations/neighbourhoods.geojson")
```  


```{r eda}
head(asheville)
dim(asheville)
names(asheville)

class(asheville_polygons)
#view(asheville_polygons)
```  

```{r prep}
asheville_rooms <- data.frame(unique(asheville$room_type))

#view(asheville_rooms)

asheville_entire.homes <- asheville %>% filter (room_type == "Entire home/apt")%>% arrange(neighbourhood)

library(sp)
coordinates(asheville_entire.homes)<- ~longitude+latitude #transform to SpatialPointsDataFrame
proj4string(asheville_entire.homes) <- CRS("+init=epsg:4326")

# listings_homes_proj <- spTransform(asheville_entire.homes, CRS("+init=epsg:4326"))
# proj4string(listings_homes_proj)


#Set polygons dataset to spatial
asheville_sp <- as(asheville_polygons, "Spatial")

#Make sure proj CRS is matching
asheville_sp_proj <- spTransform(asheville_sp, CRS(proj4string(asheville_entire.homes))) 


proj4string(asheville_sp_proj)
proj4string(asheville_entire.homes)
```  

## Asheville Airbnb Plot

```{r plots, echo=FALSE}
par(mar=c(1,1,1,1))

plot(asheville_sp)
plot(asheville_entire.homes, col="red",add=TRUE)
```  
## Spatial Interpolation  

### Proximity Polygons

```{r prox polygon}
#Proximity Polygons
require(deldir)
require(sp)

# Modified Carson Farmer code
voronoipolygons = function(layer) {
  crds <- layer@coords
  z <- deldir(crds[,1], crds[,2])
  w <- tile.list(z)
  polys <- vector(mode='list', length=length(w))
  for (i in seq(along=polys)) {
    pcrds <- cbind(w[[i]]$x, w[[i]]$y)
    pcrds <- rbind(pcrds, pcrds[1,])
    polys[[i]] <- Polygons(list(Polygon(pcrds)),
                           ID=as.character(i))
  }
  SP <- SpatialPolygons(polys)
  voronoi <- SpatialPolygonsDataFrame(SP, 
                                      data=data.frame(dummy = seq(length(SP)),
                                                      row.names=sapply(slot(SP, 'polygons'), 
                                                                       function(x) slot(x, 'ID'))))
  proj4string(voronoi) <- CRS(proj4string(layer))
  return(voronoi)
}


library(gstat)
library(tmap)
# Can't get below to work due to error.  Tried replacing x = etc.. in above function with suggested stackexchange article dummy = seq(length(SP)), but didn't work for tmap after that. (https://gis.stackexchange.com/questions/190917/r-voronoi-tesselation-from-long-lat-data)
asheville.voro <- voronoipolygons(asheville_entire.homes)

tmap_mode('plot')
fpt <- tm_shape(asheville_entire.homes) + tm_dots(size=0.1)
fvr <- tm_shape(asheville.voro) + tm_borders()
tmap_arrange(fpt,fvr)


library(gstat)
library(GISTools)
tmap_mode('view')
sh <- shading(breaks=c(50,150,250,350),
              cols=brewer.pal(5,'Purples'))
tm_shape(asheville.voro) + tm_fill(col='price',style='fixed',breaks=c(0,50,150,250,350),
                                alpha=0.6,title='Price Density')
#### Can't get map to generate.  I've tried removing every single attribute.  I've tried different versions of breaks, and I've tried changing to just tm_fill()  nothing works.

```  

### IDW

```{r idw}
library(maptools) # Required package
library(GISTools) # Required package
library(gstat) # Set up the gstat package
library(tmap)
# Define a sample grid then use it as a set of points 
# to estimate fulmar density via IDW, with alpha=1 (inverse relationship)
s.grid <- spsample(asheville.voro,type='regular',n=6000)
idw.est <- gstat::idw(price~1,asheville_entire.homes,
                      newdata=s.grid,idp=1.0)

# tmap_mode('view')
# tm_shape(idw.est) + tm_dots(col='var1.pred',border.col=NA,alpha=0.7)
# tmap_mode('view')
# idw.grid <- SpatialPixelsDataFrame(idw.est,data.frame(idw.est)) 
# tm_shape(idw.grid) + tm_raster(col='var1.pred',title='Price')
# 
# 
#How does this compare to when the idp (alpha) = 2.0
#(which indicates an inverse squared relationship)?

idw.est2 <- gstat::idw(price~1,asheville_entire.homes,
                       newdata=s.grid,idp=2.0)
idw.grid2 <- SpatialPixelsDataFrame(idw.est2,data.frame(idw.est2))
# tmap_mode('view')
# #levs <- c(0,2,4,6,8,Inf)
# tm_shape(idw.grid2) +  tm_raster(col='var1.pred',title='Price')

tmap_mode('plot')
idw1 <- tm_shape(idw.grid) +  tm_raster(col='var1.pred',title='Alpha = 1')
idw2 <- tm_shape(idw.grid2) +  tm_raster(col='var1.pred',title='Alpha = 2')
tmap_arrange(idw1,idw2)

# Extract the distinct x and y coordinates of the grid
# Extract the predicted values and form into a matrix
# of gridded values
ux <- unique(coordinates(idw.est)[,1])
uy <- unique(coordinates(idw.est)[,2])
predmat <- matrix(idw.est$var1.pred,length(ux),length(uy))

predmat2 <- matrix(idw.est2$var1.pred,length(ux),length(uy))
par(mfrow=c(1,2),mar=c(0,0,2,0))
persp(predmat,box=FALSE)
persp(predmat2,box=FALSE)
```  

### Trend Surface Analysis

```{r trend surf}
library(rgdal)
library(tmap)
library(raster)
# Create an empty grid where n is the total number of cells
grd              <- as.data.frame(spsample(asheville_entire.homes, "regular", n=50000))
names(grd)       <- c("X", "Y")
coordinates(grd) <- c("X", "Y")
gridded(grd)     <- TRUE  # Create SpatialPixel object
fullgrid(grd)    <- TRUE  # Create SpatialGrid object

# Add P's projection information to the empty grid
proj4string(grd) <- proj4string(asheville_entire.homes)

# Define the 1st order polynomial equation
f.1 <- as.formula(price ~ X + Y) 

# Add X and Y to P
asheville_entire.homes$X <- coordinates(asheville_entire.homes)[,1]
asheville_entire.homes$Y <- coordinates(asheville_entire.homes)[,2]

# Run the regression model
lm.1 <- lm( f.1, data=asheville_entire.homes)

# Use the regression model output to interpolate the surface
dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 

# Clip the interpolated raster
library(raster)
r   <- raster(dat.1st)
r.m <- mask(r, asheville_sp)

# Plot the map
tm_shape(r.m) + 
  tm_raster(n=7, palette="RdBu", 
            title="Housing Prices in Asheville") +
  tm_shape(asheville_entire.homes) + tm_dots(size=0.2) +
  tm_legend(legend.outside=TRUE)



# Define the 2nd order polynomial equation
f.2 <- as.formula(price ~ X + Y + I(X*X)+I(Y*Y) + I(X*Y))

# Add X and Y to P
asheville_entire.homes$X <- coordinates(asheville_entire.homes)[,1]
asheville_entire.homes$Y <- coordinates(asheville_entire.homes)[,2]

# Run the regression model
lm.2 <- lm( f.2, data=asheville_entire.homes)

# Use the regression model output to interpolate the surface
dat.2nd <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.2, newdata=grd))) 

# Clip the interpolated raster
r   <- raster(dat.2nd)
r.m <- mask(r, asheville_sp)

# Plot the map
tm_shape(r.m) + 
  tm_raster(n=10, palette="RdBu",
            title="Housing Prices in Asheville") +
  tm_shape(asheville_entire.homes) + tm_dots(size=0.2) +
  tm_legend(legend.outside=TRUE)
```  

## Geoweighted Regression Models 1

```{r geoweight 1}
library(GWmodel)

mean(asheville_entire.homes$price) #mean price
sd(asheville_entire.homes$price) #standard deviation

#What do the local counterparts look like?

localstats1 <- gwss(asheville_entire.homes, vars = c("price",
                                          "minimum_nights"), bw=10000) #bandwidth is 50km

head(data.frame(localstats1$SDF))

library(RColorBrewer)
plot.new()
quick.map <- function(spdf,var,legend.title,main.title) {
  x <- spdf@data[,var]
  cut.vals <- pretty(x)
  x.cut <- cut(x,cut.vals)
  cut.levels <- levels(x.cut)
  cut.band <- match(x.cut,cut.levels)
  colors <- brewer.pal(length(cut.levels),'Reds')
  par(mar=c(1,1,1,1))
  plot(ewoutline,col='grey85')
  title(main.title)
  plot(spdf,add=TRUE,col=colors[cut.band],pch=16)
  legend('topleft',cut.levels,col=colors,pch=16,bty='n',title=legend.title)
}

plot.new()
par(mfrow=c(1,2))
quick.map(localstats1$SDF, "minimum_nights",
          "100's of $", "Geographically Weighted Mean")

quick.map(localstats1$SDF, "minimum_nights",
          "100's of $", "Local Standard Deviation")
```  

## Geoweighted Regression Models 1 cont...

```{r gwr.basic 1}


gwr.res <- gwr.basic(price~minimum_nights,
                     data=asheville_entire.homes,bw=10000, kernel='gaussian')

gwr.res

gwr.res.ad <- gwr.basic(price~minimum_nights,
                        data=asheville_entire.homes,adaptive=TRUE,bw=100)
quick.map(gwr.res.ad$SDF,"minimum_nights",
          "1000's Uk Pounds per Square meter",
          "Geographically Weighted Regression Coefficient")
```  

## Geoweighted Regression Models 2

```{r geoweight 2}
library(GWmodel)

mean(asheville_entire.homes$availability_365) #mean price
sd(asheville_entire.homes$availability_365) #standard deviation

#What do the local counterparts look like?

localstats1 <- gwss(asheville_entire.homes, vars = c("price",
                                          "availability_365"), bw=10000) #bandwidth is 50km

head(data.frame(localstats1$SDF))

library(RColorBrewer)
plot.new()
quick.map <- function(spdf,var,legend.title,main.title) {
  x <- spdf@data[,var]
  cut.vals <- pretty(x)
  x.cut <- cut(x,cut.vals)
  cut.levels <- levels(x.cut)
  cut.band <- match(x.cut,cut.levels)
  colors <- brewer.pal(length(cut.levels),'Reds')
  par(mar=c(1,1,1,1))
  plot(ewoutline,col='grey85')
  title(main.title)
  plot(spdf,add=TRUE,col=colors[cut.band],pch=16)
  legend('topleft',cut.levels,col=colors,pch=16,bty='n',title=legend.title)
}

plot.new()
par(mfrow=c(1,2))
quick.map(localstats1$SDF, "availability_365",
          "100's of $", "Geographically Weighted Mean")

quick.map(localstats1$SDF, "availability_365",
          "100's of $", "Local Standard Deviation")
```  

## Geoweighted Regression Models 2 cont...

```{r gwr.basic 2}


gwr.res <- gwr.basic(price~availability_365,
                     data=asheville_entire.homes,bw=10000, kernel='gaussian')

gwr.res

gwr.res.ad <- gwr.basic(price~availability_365,
                        data=asheville_entire.homes,adaptive=TRUE,bw=100)
quick.map(gwr.res.ad$SDF,"availability_365",
          "1000's Uk Pounds per Square meter",
          "Geographically Weighted Regression Coefficient")
```